---
title: "Vast Challenge 2021 Assignment"
description: Welcome to Ong Chee Hong's Vast Challenge 2021 Assignment
author:
  - name: Ong Chee Hong
    url: https://www.linkedin.com/in/alexongch/
date: 07-04-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Vast Challenge 2021: Mini-Challenge 2

# Background
Many of the Abila, Kronos-based employees of GAStech have company cars which are approved for both personal and business use. Those who do not have company cars have the ability to check out company trucks for business use, but these trucks cannot be used for personal business.

Employees with company cars are happy to have these vehicles, because the company cars are generally much higher quality than the cars they would be able to afford otherwise. However, GAStech does not trust their employees. Without the employees? knowledge, GAStech has installed geospatial tracking software in the company vehicles. The vehicles are tracked periodically as long as they are moving.

This vehicle tracking data has been made available to law enforcement to support their investigation. Unfortunately, data is not available for the day the GAStech employees went missing. Data is only available for the two weeks prior to the disappearance.

To promote local businesses, Kronos based companies provide a Kronos Kares benefit card to GASTech employees giving them discounts and rewards in exchange for collecting information about their credit card purchases and preferences as recorded on loyalty cards. This data has been made available to investigators in the hopes that it can help resolve the situation. However, Kronos Kares does not collect personal information beyond purchases.

As a visual analytics expert assisting law enforcement, your mission is to identify which GASTech employees made which purchases and identify suspicious patterns of behavior. You must cope with uncertainties that result from missing, conflicting, and imperfect data to make recommendations for further investigation.


# Questions
1.	Using just the credit and loyalty card data, identify the most popular locations, and when they are popular. What anomalies do you see? What corrections would you recommend to correct these anomalies? Please limit your answer to 8 images and 300 words.

2.	Add the vehicle data to your analysis of the credit and loyalty card data. How does your assessment of the anomalies in question 1 change based on this new data? What discrepancies between vehicle, credit, and loyalty card data do you find? Please limit your answer to 8 images and 500 words.

3.	Can you infer the owners of each credit card and loyalty card? What is your evidence? Where are there uncertainties in your method? Where are there uncertainties in the data? Please limit your answer to 8 images and 500 words.

4.	Given the data sources provided, identify potential informal or unofficial relationships among GASTech personnel. Provide evidence for these relationships. Please limit your response to 8 images and 500 words.

5.	Do you see evidence of suspicious activity? Identify 1- 10 locations where you believe the suspicious activity is occurring, and why Please limit your response to 10 images and 500 words.

6.	If you solved this mini-challenge in 2014, how did you approach it differently this year?

# Installing required R pacakages
```{r}
packages = c('DT','ggiraph','plotly','tidyverse', 'raster','sf','clock','tmap', 'rgdal','dplyr', 'tidyr', 'textclean', "plotly", "forcats")
for(p in packages){
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Provided data and information
There are 3 different types of data & information provided. A geospatial dataset, csv files and a jpg file consisting of the Abila tourist map.

Below is the information of all the data provided.

### 1. A list of vehicle assignments by employee, in CSV format (car-assignments.csv)

1.1 Employee Last Name

1.2	Employee First Name

1.3	Car ID (integer)

1.4	Current Employment Type (Department; categorical)

1.5	Current Employment Title (job title; categorical)

```{r}
vehicle <- read_csv("data/mc2/car-assignments.csv")

DT::datatable(vehicle)

```

### 2.	ESRI shapefiles of Abila and Kronos (in the Geospatial folder)

### 3.	A CSV file of vehicle tracking data 

3.1	Timestamp

3.2	Car ID (integer)

3.3	Latitude

3.4	Longitude

### 4.	A CSV file containing loyalty card transaction data (loyalty_data.csv)

4.1	Timestamp

4.2	Location (name of the business)

4.3	Price (real)

4.4	Loyalty Number (A 5-character code starting with L that is unique for each card)

### 5.	A CSV file containing credit and debit card transaction data (cc_data.csv)

5.1	Timestamp

5.2	Location (name of the business)

5.3	Price (real)

5.4	Last 4 digits of the credit or debit card number

### 6.	A tourist map of Abila with locations of interest identified, in JPEG format (MC2-Tourist.jpg)


# Literature Review

This year Vast Challenge uses the same question based on 2014 Vast Challenge. As such, we will conduct literature review on past 2014 Vast Challenge submission based on MC2 – Patterns of Life Analysis. We will select some of the past submission to identify some of the gaps that can potentially utilized interactive data visualization techniques to improve user experiences. The repository for 2014 Vast Challenge – MC2: Patterns of Life Analysis submissions can be found  [here](http://visualdata.wustl.edu/varepository/VAST%20Challenge%202014/challenges/MC2%20-%20Patterns%20of%20Life%20Analysis/)

#Importing of data

```{r}
vehicle <- read_csv("data/mc2/car-assignments.csv")
cc <- read_csv("data/mc2/cc_data.csv", locale = locale(encoding = "ASCII"))
gps <- read_csv("data/mc2/gps.csv")
loyalty <- read_csv("data/mc2/loyalty_data.csv")
```

```{r}
guess_encoding(cc)
```

#Answers

### 1.  Using just the credit and loyalty card data, identify the most popular locations, and when they are popular. What anomalies do you see? What corrections would you recommend to correct these anomalies? Please limit your answer to 8 images and 300 words.

#### 1.1 Examining both credit and loyalty card data

```{r}

glimpse(cc)
glimpse(loyalty)
```
From both the dataset. We are able to see that the credit card dataset has 98 rows more than the loyalty dataset. Based on the given information, the count of credit card purchases should be the same as the count of loyalty cards as both data should be recorded. One assumption to this anomaly might be that those purchases without loyalty data were employees that made purchases without using the Kronos Kares benefit card for unknown reason.


Next, we will change the datatype of last4ccnum and loyaltynum to categorical data and the timestamp for both cc and loyalty data.

```{r}
cc$last4ccnum <- as.factor(cc$last4ccnum)
loyalty$loyaltynum <- as.factor(loyalty$loyaltynum)

cc$timestamp <- date_time_parse(cc$timestamp,
                                 zone = "",
                                 format = "%m/%d/%Y %H:%M")

loyalty$timestamp <- date_parse(loyalty$timestamp,
                                 format = "%m/%d/%Y")

```
```{r}
glimpse(cc)
glimpse(loyalty)

```


split into day, date and time 
```{r}
cc_dtsplit <- cc %>%
  mutate(day = date_weekday_factor(cc$timestamp), date =  as_date(cc$timestamp), time = format(cc$timestamp, format = "%H:%M"))

cc_dtsplit

```

```{r}
replace_non_ascii(cc_dtsplit['location'],'',remove.nonconverted = TRUE)

loyalty_recode <- mutate(loyalty, location = recode(location, "Katerina’s Café" = "Katerina's Cafe"))

cc_dtsplit

```



rename loyalty's timestamp
```{r}
loyalty_dt <- rename(loyalty, date = timestamp)


```


```{r}
glimpse(cc_dtsplit)
glimpse(loyalty_dt)
```
making time into groups of 15 min - optional
```{r}
#cc_modified <- cc_recode %>%
  #mutate(time_15 = cut(timestamp, breaks="15 min"))

```



inner join
```{r}
cc_loyalty_join <- cc_dtsplit %>%
  inner_join(loyalty_dt, by = c("date","location", "price"))

cc_loyalty_join
```
Notice that after we have inner join both the cleaned dataset. There is only 932 rows while the orignal cc_dtsplit and has 1490 and loyalty_dt has 1392. This anomalies suggested that some of the people that uses their credit card did not use their loyalty card during purchases.


Before we begin further analysis on further anomalies between the two datasets. We shall conduct some exploratory data analysis. First, we will analyse the most popular place visited during the duration of study.

```{r}
cc_location_count <- cc_loyalty_join %>%
  group_by(location) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

cc_location_count

```

```{r}
cc_join <- cc_loyalty_join %>%
  count(location) %>%
  mutate(location = fct_reorder(location, n, .desc =TRUE)) %>%
  plot_ly(x = ~location, y = ~n, marker = list(color = ~n)) %>%
  add_bars(name = "inner-join") %>%
  layout(title = "Total number of visitation by area", xaxis = list(title = ""),yaxis = list(title = "Number of visitors"))

cc_join
```

From the chart above, we can observe that "Guy's Gyros", "Hippokampos", "Brew've Been Serve" have the most vistitation during the period of study.




```{r}
cc_loyalty_group <- cc_loyalty_join %>%
  group_by(location,date,day,time) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

cc_loyalty_group

```
```{r}
cc_loyalty_day <- cc_loyalty_join %>%
  unite(col = "daytime", day,time, sep = " ",  remove =FALSE) %>%
  group_by(location, daytime) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

cc_loyalty_day
```
```{r}
cc_loyalty_day %>%
  filter(count >=5) %>%
  plot_ly(x = ~daytime, y = ~count, color = ~location) %>%
  add_bars() %>%
  layout (title = "Total counts of visitors by day and time", xaxis = list(title = "", size = 3, dtick = ~daytime), yaxis = list(title = "Number of visitors") )

```


```{r}
cc_loyalty_top4 <- cc_loyalty_group %>%
  filter(count >= 3) %>%
  ungroup()

```


```{r}
cc_loyalty_filter <- cc_loyalty_group %>%
  filter(location %in% c("Bean There Done That", "Guy's Gyros", "Jack's Magical Beans"))

cc_loyalty_filter
```

```{r}
cc_loyalty_top4 %>%
  plot_ly(x = ~date, y =~count, color = ~location, hoverinfo = "text",
          text = ~paste("Location:", location, "<br>","Date:", date, "<br>","Time:", time, "<br>", "Day:", day)) %>%
  add_lines() %>%
  add_markers(showlegend = FALSE) %>%
  layout(title = "Most frequented places by groups of people", xaxis = list(title = "", dtick = ~day), yaxis = list(title = "Number of people"))
```

anti_join
```{r}
cc_loyalty_antijoin <- cc_dtsplit %>%
  anti_join(loyalty_dt, by = c("date","location", "price"))

cc_loyalty_join
```

```{r}
cc_anti <- cc_loyalty_antijoin %>%
  count(location) %>%
  mutate(location = fct_reorder(location, n, .desc =TRUE)) %>%
  plot_ly(x = ~location, y = ~n, marker = list(color = ~n)) %>%
  add_bars(name = "anti-join") %>%
  layout(title = "Total number of visitation by area", xaxis = list(title = ""),yaxis = list(title = "Number of visitors"))

cc_anti
```

```{r}
subplot(cc_join, cc_anti, nrows = 1, shareX = TRUE, shareY = TRUE)

```

```{r}
cc_loyalty_day_anti <- cc_loyalty_antijoin %>%
  unite(col = "daytime", day,time, sep = " ",  remove =FALSE) %>%
  group_by(location, daytime) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

cc_loyalty_day_anti
```

```{r}
cc_loyalty_day_anti %>%
  filter(count >=3) %>%
  plot_ly(x = ~daytime, y = ~count, color = ~location) %>%
  add_bars() %>%
  layout (title = "Total counts of visitors by day and time", xaxis = list(title = "", size = 3, dtick = ~daytime), yaxis = list(title = "Number of visitors") )

```

```{r}
cc_loyalty_anti_group <- cc_loyalty_antijoin %>%
  group_by(location, day, date, time) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  ungroup()

cc_loyalty_anti_group
```

```{r}
cc_loyalty_anti_group %>%
  filter(count >= 2) %>%
  plot_ly(x = ~date, y =~count, color = ~location, hoverinfo = "text",
          text = ~paste("Location:", location, "<br>","Date:", date, "<br>","Time:", time, "<br>", "Day:", day)) %>%
  add_lines() %>%
  add_markers(showlegend = FALSE) %>%
  layout(title = "Most frequented places by groups of people", xaxis = list(title = "", dtick = ~date), yaxis = list(title = "Number of people"))
```
